---
title: "Brundle Example"
author: "Andrew Holding"
date: "8/8/2017"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Brundle Examples

This markdown provides an example of a workflow using Brundle applied to a minmal 
dataset as included in the BrudleData package. 

The packages are found on GitHub as AndrewHolding/Brundle & 
AndrewHolding/BrundleData. They can be installed with the following code.

```{r eval=FALSE}

install.packages("devtools")
library(devtools)

install_github("AndrewHolding/packagename")

```

Once installed we do not need to install them again and can load them as normal.

```{r, message=FALSE, warning=FALSE}
library(Brundle)
library(BrundleData)

```

The intial steps of the Brundle Pipleline are to set the varibles. Here we are
using the data from the BundleData package which contains two samplesheets 
formatted as required by DiffBind. They both refer to the same data, but one
provides BED files for only the CTCF peaks while the provides BED files for only
the ER regions. The CTCF regions are to provide our control peaks, while the ER
binding provides our experimental peaks changes. In this example we have treated
MCF7 cells with Fulvestrant.

```{r}

#Set up the intial varible
jg.controlMinOverlap      <- 5
jg.controlSampleSheet     <-
    system.file("extdata", "samplesheet_SLX14438_hs_CTCF_DBA.csv", package =
                    "Brundle")
jg.experimentSampleSheet  <-
    system.file("extdata", "samplesheet_SLX14438_hs_ER_DBA.csv", package =
                    "Brundle")
jg.treatedCondition       =  "Fulvestrant"
jg.untreatedCondition     =  "none"

```

Once configured we load the data from the samples sheets as normal with
DiffBind. This provides us with two DiffBind objects. One Experimental and one
Control.

```{r}

setwd(system.file("extdata",package="BrundleData"))

dbaExperiment <- jg.getDba(jg.experimentSampleSheet, bRemoveDuplicates=TRUE)
dbaControl    <- jg.getDba(jg.controlSampleSheet, bRemoveDuplicates=TRUE)

```

We then use Brundle to extract the data from the DiffBind object to generate
a peakset. This provides us with the read count at each peak location for each
sample.

```{r}
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
```

To normalise the data we need to counts of the control and treated samples
seperately. This uses the original information we provided at the start of the
script to split the control samples into two matrices. For convience we also 
record the names of the samples relating to each condition.

```{r}

setwd(system.file("extdata",package="BrundleData"))

#Get counts for the treated control samples.
jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset,
                                             jg.controlSampleSheet,
                                             jg.treatedCondition )
#Repeat for the untreated/control samples
jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
                                               jg.controlSampleSheet,
                                               jg.untreatedCondition)

#Get the sample names for replicates that represent the two conditions.
jg.untreatedNames <- names(jg.controlCountsUntreated)
jg.treatedNames   <- names(jg.controlCountsTreated)
```

Next we generate a normalization coefficient from the data. Typically this is
visualized with the included plot function but the step is not required, it can
be calculated directly from the data.

```{r}
jg.plotNormalization(jg.controlCountsTreated,
                     jg.controlCountsUntreated)

#Calculate the noralisation coefficent
jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
                                               jg.controlCountsUntreated)
```

To reinsert the data into DiffBind we calculate a correction factor. This is
essential as DiffBind will try to normalise our data, this correction factor 
ensures that our normalization coefficent is applied correctly.

```{r}

setwd(system.file("extdata",package="BrundleData"))

jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
                                            jg.treatedNames,
                                            jg.untreatedNames)
```

We then apply the normalisation coefficent and correction factor to the 
treated samples.

```{r}
jg.experimentPeaksetNormalised<-jg.applyNormalisation(jg.experimentPeakset,
                                                      jg.coefficient,
                                                      jg.correctionFactor,
                                                      jg.treatedNames)
```

For convience the we return the data to DiffBind (using a modified version from
https://github.com/andrewholding/flypeaks/tree/master/Diffbind) and use DiffBind 
to analyze the data. We could then go on to generate a DiffBind report. As this 
is the analysis of only chromosome 22 we get only a small number of 
differentially bound sites, nonetheless, the proceedure documented here will 
work for much larger datasets.

```{r}
jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
                                    jg.experimentPeaksetNormalised)

dba.analysis<-dba.analyze(jg.dba)
dba.plotMA(dba.analysis,bSmooth=FALSE,bFlip = TRUE)
```

